---
title: "Grain–Age Key Estimation Tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Grain–Age Key Estimation Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
library(grainkey)
library(ggplot2)
library(magrittr)  # Para usar %>% en la viñeta

```

# Introduction

This vignette demonstrates how to estimate and visualize the age composition para el caso base of anchovy (*Engraulis encrasicolus*) using the grainkey package. The package allows users to build a grain–age key based on biological samples and explore commercial catch composition by grain class.

## Estimating the Grain–Age Key from Biological Samples

### Load Example Data

```{r}
data(Database)
head(Database)
```


### Build the Grain–Age Key

Use `build_grain_key()` to calculate the age composition and weight-at-age by grain class.

```{r}
result <- build_grain_key(Database)
```

## Viewing the Age Composition Table

```{r}
result$grain_key # Proportion-at-age

result$weight_key # Mean weight-at-age
```

## Visualizing the Grain–Age Key


### Age Distribution by Grain Class
```{r}
plot_grain_key(result, plot_type="percentage")
```

### Mean Weight-at-Age by Grain Class
```{r}
plot_grain_key(result, plot_type="mean_weight")
```


## Estimating Catch Composition from Sales Data

### Load Simulated Sales Data
```{r}
data("data_NV")

notaV<-data_NV
```


### Estimate Catch by Grain Class
```{r}
res_default     <- calc_grain_catch(notaV)
res_t1_to_t3    <- calc_grain_catch(notaV, reassign = TRUE, talla1_to = "TALLA 3")
res_no_reassign <- calc_grain_catch(notaV, reassign = FALSE)
```

### Visualize Catch by Grain Class and Quarter
```{r}
plot_grain_catch(res_no_reassign)
plot_grain_catch(res_default)
plot_grain_catch(res_t1_to_t3)
```

## Estimating Catch-at-Age

```{r}
catch_by_age <- calc_catch_by_age(result$grain_key, res_default)
catch_by_age
```

### Plot Catch-at-Age

```{r}
# Graficar
plot_catch_by_age(catch_by_age)
```

# Conclusion

The grainkey package provides tools for translating biological and commercial data into grain–age keys and applying them to catch records to support age-based catch estimation for fisheries assessment.

